### 2019-12-31 2019_edits/R/fecundity_code/egg_models_2.Rmd
big backup

---
title: "August-ish Manuscript Edits"
author: "Adam A Kemberling"
date: "10/1/2019"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
    mathjax: null
editor_options: 
  chunk_output_type: console
---

# Crab Mass x Carapace Width

```{r, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      comment = NA, 
                      message = FALSE, 
                      warning = FALSE)
```

```{r}
#Load Packages
library(tidyverse, quietly = TRUE)
library(broom)
library(knitr)
library(here)
library(rnaturalearth)
library(sf)
library(scales)
library(magrittr)
library(statmod)

#Spatial Polygons
usmap <- ne_states(country = "united states of america") %>% 
  st_as_sf() %>% 
  filter(region == "South")
mex <- ne_states(country = "mexico") %>% 
  st_as_sf()

#Load Data

#Egg Data
eggdat <- read_csv(here("data/repo/seamap_summerandfall2017_crabs.csv"), guess_max = 10000, col_types = cols())
eggdat <- eggdat %>% 
  mutate(Egg_stage = factor(eggdat$Egg_stage),
         molt_stage = factor(molt_stage),
         Estimated_fecundity = ifelse(Estimated_fecundity == "#DIV/0!", NA, Estimated_fecundity),
         Estimated_fecundity = as.numeric(Estimated_fecundity))

#Stations
stat_dat <- read_csv(here("data/repo/2017_allcrabs_onerow_withstationdata.csv"), guess_max = 10000, col_types = cols())
eggdat <- stat_dat %>%  
  select(Unique_ID, Agency, STA_TIME, END_TIME,STA_LON, STA_LAT) %>% 
  right_join(eggdat, by = "Unique_ID")

#Regions and Statzones
# Add the statzone using statzone polygon - use this to make regions consistent between models
szones <- read_sf(here("data/gis_files","NMFS_Fishing_Grids_GOM_2013.shp"), crs = 4326)

# Add a better StatZone ID by spatial overlap with NMFS Zones
eggdat <- eggdat %>% 
  st_as_sf(coords = c("STA_LON", "STA_LAT"), crs = 4326, remove = FALSE) %>% 
  st_join(szones["StatZone"]) %>% 
  st_set_geometry(NULL) %>% 
  as.data.frame() %>% 
  mutate(StatZone = factor(StatZone),
         StatZone = fct_rev(StatZone))

#Regions That are less arbitrary
eggdat <- eggdat %>% 
  mutate(region = ifelse(StatZone %in% seq("21", "18"), "Texas", "Louisiana"),
         region = ifelse(StatZone %in% seq("12", "10"), "MS Bight", region),
         region = ifelse(StatZone %in% seq("1", "9"), "Florida", region),
         region = factor(region),
         #Make bins for egg stage
         es_bins = ifelse(Egg_stage %in% c(1,3), "early", "late"),
         es_bins = ifelse(Egg_stage %in% c(4:6), "middle", es_bins),
         es_bins = factor(es_bins, levels = c("early", "middle", "late"))
  )

#Fix egg volume issue
eggdat <- eggdat %>% 
  mutate(new_vol = (4/3) * pi * ((diam1_mm + diam2_mm) / 4) ^ 3) %>%
  filter(is.na(Egg_stage) == FALSE)

#Write egg data out as final analysis set
write_csv(eggdat, 
          here::here("data", "repo", "analysis_sets", "seamap_repro_eggs.csv"), 
          col_names = TRUE)

#Make a separate DF for just the ovigerous crabs
# #Take the mean value of each egg measurement so theres one value per crab
fecunddat_means <- eggdat %>%
  filter(ovigerous == "yes") %>%
  group_by(Unique_ID) %>%
  summarise(Estimated_fecundity = mean(Estimated_fecundity, na.rm = T),
            egg_volume = mean(new_vol, na.rm = T)) %>%
  ungroup()

#Pull crab specific details and combine
fecunddat <- eggdat %>% 
  filter(ovigerous == "yes") %>% 
  distinct(Unique_ID, .keep_all = TRUE) %>% 
  select(Unique_ID, ovigerous, STA_LON, STA_LAT, Season, region, Crab_weight, Carapace_width, Standard_cw, molt_stage, Eggmass_weight, Egg_stage, Percent_fert, es_bins) %>% 
    left_join(fecunddat_means, by = "Unique_ID")

#Write that out for crab specific means
write_csv(fecunddat, 
          here::here("data", "repo", "analysis_sets", "seamap_repro_eggmass_avg.csv"), 
          col_names = TRUE)

```

## Total Carapace Width

```{r crab condition}

#Total Carapace Width
model_corr <- fecunddat %>% filter(is.na(Carapace_width) == FALSE, is.na(Crab_weight) == FALSE) %$% 
  cor(.$Carapace_width, .$Crab_weight) %>% round(3)

fecunddat %>% 
  ggplot(aes(Carapace_width, Crab_weight)) + 
  geom_point() +
  geom_smooth(method = "glm",
              formula = y~x,
              method.args = list(family = gaussian(link = 'identity')),
              color = "black", linetype = 2) +
  labs(x = "Total Carapace Width", 
       y = "Total Crab Mass (g)") +
  annotate(geom = "text", x = 130, y = 300, label = str_c("Correllation: \n", model_corr))
```

### Fits - Total CW

```{r}
t_cw_mod <- glm(Crab_weight ~ Carapace_width, 
                data = fecunddat, 
                family = gaussian(link = "identity"))

t1 <- qresid(t_cw_mod)
qqnorm(t1)
abline(0, 1, col = "red")

hist(t_cw_mod$residuals, main = "Total Carapace Width Residuals")

```

## Standard Carapace Width

```{r}

# Standard Carapace Width
model_corr <- fecunddat %>% filter(is.na(Standard_cw) == FALSE, is.na(Crab_weight) == FALSE) %$% 
  cor(.$Standard_cw, .$Crab_weight) %>% round(3)

fecunddat %>% 
  ggplot(aes(Standard_cw, Crab_weight)) + 
  geom_point() +
  geom_smooth(method = "glm",
              formula = y~x,
              method.args = list(family = gaussian(link = 'identity')),
              color = "black", linetype = 2) +
  labs(x = "Standard Carapace Width", y = "Total Crab Mass (g)") +
  annotate(geom = "text", x = 100, y = 300, label = str_c("Correllation: \n", model_corr))
```

### Fits - Standard CW

```{r}
s_cw_mod <- glm(Crab_weight ~ Standard_cw, 
                data = fecunddat, 
                family = gaussian(link = "identity"))

t1 <- qresid(s_cw_mod)
qqnorm(t1)
abline(0, 1, col = "red")

hist(s_cw_mod$residuals, main = "Standard Carapace Width Residuals")
```

# Fecundity x crab mass

```{r}

# Standard Carapace Width
model_corr <- fecunddat %>% 
  filter(is.na(Crab_weight) == FALSE, is.na(Estimated_fecundity) == FALSE) %$% 
  cor(.$Crab_weight, .$Estimated_fecundity) %>% round(3)

fecunddat %>% 
  ggplot(aes(Crab_weight, Estimated_fecundity)) + 
  geom_point() +
  geom_smooth(method = "glm",
              formula = y~x,
              method.args = list(family = gaussian(link = 'identity')),
              color = "black", linetype = 2) +
  labs(x = "Crab Weight (w/o eggmass in g)", y = "Estimated Fecundity (Total Eggs)") +
  scale_y_continuous(labels = comma_format()) +
  annotate(geom = "text", x = 130, y = 6250000, label = str_c("Correllation: \n", model_corr))

```

### Fits

```{r}
fecund_weight_mod <- glm(Estimated_fecundity ~ Crab_weight, 
                data = fecunddat, 
                family = gaussian(link = "identity"))

t1 <- qresid(fecund_weight_mod)
qqnorm(t1)
abline(0, 1, col = "red")

hist(fecund_weight_mod$residuals, main = "Fecundity ~ Crab Weight - Residuals")
```

# Fecundity x Carapace Width

## Standard Identity link
```{r, eval = FALSE}

plot_mod <- glm(data = fecunddat, Estimated_fecundity ~ Carapace_width, family = gaussian(link = "identity"))
summary(plot_mod)
```

```{r}
fecunddat %>% 
  ggplot(aes(Carapace_width, Estimated_fecundity)) + 
  geom_point() +
  geom_smooth(method = "glm",
              formula = y~x,
              method.args = list(family = gaussian(link = 'identity')),
              color = "black", linetype = 2) +
  labs(x = "Total Carapace Width (mm)", y = "Estimated Fecundity (Total Eggs)") +
  scale_y_continuous(labels = comma_format()) 
```

### Fits

```{r}
plot_mod <- glm(data = fecunddat, Estimated_fecundity ~ Carapace_width, family = gaussian(link = "identity"))
t1 <- qresid(plot_mod)
qqnorm(t1)
abline(0, 1, col = "red")

hist(plot_mod$residuals, main = "Identity-link residuals")

```

## Log-link

```{r, eval = FALSE}

plot_mod <- glm(data = fecunddat, Estimated_fecundity ~ Carapace_width, family = gaussian(link = "log"))
summary(plot_mod)
```

```{r}
fecunddat %>% 
  ggplot(aes(Carapace_width, Estimated_fecundity)) + 
  geom_point() +
  geom_smooth(method = "glm",
              formula = y~x,
              method.args = list(family = gaussian(link = 'log')),
              color = "black", linetype = 2) +
  labs(x = "Total Carapace Width (mm)", y = "Estimated Fecundity (Total Eggs)") +
  scale_y_continuous(labels = comma_format()) 
```

### Fits

```{r}
plot_mod <- glm(data = fecunddat, Estimated_fecundity ~ Carapace_width, family = gaussian(link = "log"))
t1 <- qresid(plot_mod)
qqnorm(t1)
abline(0, 1, col = "red")

hist(plot_mod$residuals, main = "log-link residuals")

```

# Egg Size x Body Size * Developmental Bins

```{r egg size and carapace width}

library(ggpirate)
fecunddat %>% 
  ggplot(aes(x = es_bins, y = egg_volume, color = egg_volume, fill = egg_volume)) +
  geom_pirate(bars = FALSE) +
  labs(
    x = "Egg Development Stage",
    y = "Egg Volume"
  )
```

### Anova - Egg Development

```{r}
anova_mod <- lm(egg_volume ~ es_bins, data = fecunddat)
anova(anova_mod)
```