### 2019-12-31 2019_edits/R/fecundity_code/eggmodels_markdown.Rmd

---
title: "Egg Analyses Model Markup"
author: "Adam A Kemberling"
date: "2/4/2019"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
    mathjax: null
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = NA, message = FALSE, warning = FALSE)

library(tidyverse, quietly = TRUE)
library(broom)
library(knitr)
library(here)
library(rnaturalearth)
library(sf)
library(scales)

#Mixed Modeling PAckages
library(lme4)
library(afex)
library(broom.mixed)
library(emmeans)

## squash panels together
zero_margin <- theme(panel.spacing=grid::unit(0,"lines")) 
library(lattice)
library(dotwhisker)  ## coefficient plots
library(broom.mixed)
conflicted::conflict_prefer("select", "dplyr")
conflicted::conflict_prefer("filter", "dplyr")
conflicted::conflict_prefer("lmer", "lme4")

#Format ggplot for rest of document
theme_set(theme_bw() + theme(panel.grid.major = element_blank(), 
                             panel.grid.minor = element_blank(), 
                             panel.border = element_rect(fill = NA)))

```

# Load and clean data

```{r}
#crabdat <- read_csv("~/Dropbox (The Craboratory)/The Craboratory/Kemberling/SEAMAP_2017/2017_allcrabs_onerow_withstationdata.csv",
# guess_max = 10000, col_types = cols()) 
# eggdat <- read_csv("~/Dropbox (The Craboratory)/The Craboratory/Kemberling/SEAMAP_2017/Reproductive_analysis/seamap_summerandfall2017_crabs.csv", guess_max = 10000, col_types = cols())

crabdat <- read_csv(here("data/repo/2017_allcrabs_onerow_withstationdata.csv"), guess_max = 10000, col_types = cols())
eggdat <- read_csv(here("data/repo/seamap_summerandfall2017_crabs.csv"), guess_max = 10000, col_types = cols())

eggdat <- eggdat %>% 
  mutate(Egg_stage = factor(eggdat$Egg_stage),
         molt_stage = factor(molt_stage),
         Estimated_fecundity = ifelse(Estimated_fecundity == "#DIV/0!", NA, Estimated_fecundity),
         Estimated_fecundity = as.numeric(Estimated_fecundity))

#Zack wants to see if regions matter, so lets add them in now, gonna need the station info data though
#stat_dat <- read_csv("~/Dropbox (The Craboratory)/The Craboratory/Kemberling/SEAMAP_2017/2017_allcrabs_onerow_withstationdata.csv", guess_max = 10000, col_types = cols())

stat_dat <- read_csv(here("data/repo/2017_allcrabs_onerow_withstationdata.csv"), guess_max = 10000, col_types = cols())

# #verify that the unique ID's are the same, they were made in a separate script
# stat_dat$Unique_ID %in% eggdat$Unique_ID
# eggdat$Unique_ID %in% stat_dat$Unique_ID

#Pull the columns we want out, merge them to eggdat by the station info
eggdat <- stat_dat %>%  
  select(Unique_ID, Agency, STA_TIME, END_TIME,STA_LON, STA_LAT) %>% 
  right_join(eggdat, by = "Unique_ID")

```

# Add Regions + Explore Data {.tabset .tabset-fade .tabset-pills}

Regions:

 * S. Texas (Matagorda Bay south)     
 * TX/LA: (between Matagorda Bay and MS River delta)      
 * MS Bight: LA east of MS River, MS, AL     
 * Florida

```{r}

#basemap for reference
#usmap <- map_data("state")

usmap <- ne_states(country = "united states of america") %>% 
  st_as_sf() %>% 
  filter(region == "South")

mex <- ne_states(country = "mexico") %>% 
  st_as_sf()

# Add the statzone using statzone polygon - use this to make regions consistent between models
#szones <- read_sf(here("GIS_objects",  "NMFS_Fishing_Grids_GOM_2013", "NMFS_Fishing_Grids_GOM_2013.shp"), crs = 4326)
szones <- read_sf(here("data/gis_files","NMFS_Fishing_Grids_GOM_2013.shp"), crs = 4326)

# Add a better StatZone ID by spatial overlap with NMFS Zones
eggdat <- eggdat %>% 
  st_as_sf(coords = c("STA_LON", "STA_LAT"), crs = 4326, remove = FALSE) %>% 
  st_join(szones["StatZone"]) %>% 
  st_set_geometry(NULL) %>% 
  as.data.frame() %>% 
  mutate(StatZone = factor(StatZone),
         StatZone = fct_rev(StatZone))
```

## Map Stations with Repro Data

```{r}
# #show where they are
# ggplot() + 
#   geom_point(data = eggdat, aes(STA_LON, STA_LAT)) + 
#   geom_sf(data = usmap, fill = "antiquewhite1") + 
#   geom_sf(data = mex, fill = "antiquewhite1") +
#   coord_sf(xlim = c(-80.5, -97.5), ylim = c(24.75, 31)) +
#   labs(x= "", y = "") +
#   theme_minimal() +
#   ggtitle("Raw Data")

# #Make regions
# eggdat <- eggdat %>% 
#   mutate(region = ifelse(STA_LAT < 28.387595 & STA_LON < -96, "South Texas", "TX/LA"),
#          region = ifelse(STA_LON > -89.2, "MS Bight", region),
#          region = ifelse(STA_LON > -87.5, "FL", region),
#          region = factor(region))
# 
# eggdat %>% count(region, StatZone)
# 
# 
# 
# #check by plotting
# ggplot() + 
#   geom_point(data = eggdat, aes(STA_LON, STA_LAT, color = region)) + 
#   geom_sf(data = usmap, fill = "antiquewhite1") + 
#   geom_sf(data = mex, fill = "antiquewhite1") +
#   coord_sf(xlim = c(-80.5, -97.5), ylim = c(24.75, 31)) +
#   labs(x= "", y = "") +
#   theme_minimal() + theme(legend.position = "bottom",
#                           panel.border = element_rect(fill = NA)) +
#   ggtitle("Color by Region")

#Regions That are less arbitrary
eggdat <- eggdat %>% 
  mutate(region = ifelse(StatZone %in% seq("21", "18"), "Texas", "Louisiana"),
         region = ifelse(StatZone %in% seq("12", "10"), "MS Bight", region),
         region = ifelse(StatZone %in% seq("1", "9"), "Florida", region),
         region = factor(region))

#check by plotting
ggplot() + 
  geom_point(data = eggdat, aes(STA_LON, STA_LAT, color = StatZone, shape = region)) + 
  geom_sf(data = usmap, fill = "antiquewhite1") + 
  geom_sf(data = mex, fill = "antiquewhite1") +
  coord_sf(xlim = c(-80.5, -97.5), ylim = c(24.75, 31)) +
  labs(x= "", y = "") +
  theme_minimal() + theme(legend.position = "bottom",
                          panel.border = element_rect(fill = NA)) +
  guides(shape = guide_legend(title = "Region", title.position = "top", title.hjust = 0.5),
         color = guide_legend(title = "Stat Zone", title.position = "top", title.hjust = 0.5)) +
  ggtitle("Sample Regions and Corresponding Stat Zones")

```

***

##  Fix Egg Volume Measurement

I noticed a few crabs had egg sizes that were much higher than others, so going to recalculate the volume from the egg dimensions to compare.

```{r}
#Looks like the calculated volumes might be deviating from 4/3*pi*r^3, so I'm gonna check that here
eggdat <- eggdat %>% 
  mutate(new_vol = (4/3) * pi * ((diam1_mm + diam2_mm) / 4) ^ 3) %>%
  filter(is.na(Egg_stage) == FALSE)

p1 <- eggdat %>% 
  ggplot() + 
  geom_point(aes(egg_volume, new_vol, color = Unique_ID), show.legend = FALSE) +
  ggtitle("Raw volume vs. Corrected")
p1
#plotly::ggplotly(p1)

```

So the excel file on the dropbox has some inconsistencies with the estimated fecundity, and other columns that were drag-down in excel.

The values that appeared much higher have the correct volume calculations, the others have something different.

finish analysis with new_vol.

**Pull Crab Specific Fecundity Details: **

For analyses relating to egg characteristics (egg volume, estimated fecundity), measurements will be the mean value for each crab to avoid dealing with repeated measures.

```{r pull fecundity data}

# #Check to see if hand calculations (Estimated_fecundity) = programatic estimates (mean_fecundity)
# eggdat %>% 
#   filter(ovigerous == "yes") %>% 
#   group_by(Unique_ID) %>% 
#   summarise(mean_fecundity = mean(Estimated_fecundity, na.rm = TRUE),
#             mean_volume    = mean(new_vol, na.rm = TRUE)) %>% 
#   right_join(eggdat, by = "Unique_ID") %>% 
#   select(Unique_ID, Estimated_fecundity, mean_fecundity)

# # We cool, just use Estimated_fecundity, but pull one entry for each crab

# #Take the mean value of each egg measurement so theres one value per crab
fecunddat_means <- eggdat %>%
  filter(ovigerous == "yes") %>%
  group_by(Unique_ID) %>%
  summarise(Estimated_fecundity = mean(Estimated_fecundity, na.rm = T),
            egg_volume = mean(new_vol, na.rm = T)) %>%
  ungroup()

#Pull crab specific details and combine
fecunddat <- eggdat %>% 
  filter(ovigerous == "yes") %>% 
  distinct(Unique_ID, .keep_all = TRUE) %>% 
  select(Unique_ID, ovigerous, STA_LON, STA_LAT, Season, region, Crab_weight, Carapace_width, Standard_cw,                   molt_stage, Eggmass_weight, Egg_stage, Percent_fert) %>%

    left_join(fecunddat_means, by = "Unique_ID")  %>%
  mutate(
    #Make bins for egg stage
    es_bins = ifelse(Egg_stage %in% c(1,3), "early", "late"),
    es_bins = ifelse(Egg_stage %in% c(4:6), "middle", es_bins),
    es_bins = factor(es_bins, levels = c("early", "middle", "late"))
  )

#Fecundity Summary Stats
fecundity_max <-  round(max(fecunddat$Estimated_fecundity) / 1e6, 3)
fecundity_min <-  round(min(fecunddat$Estimated_fecundity) / 1e6, 3)
fecundity_mean <- round(mean(fecunddat$Estimated_fecundity) / 1e6, 3)
fecundity_se <-   round( 
  (sd(fecunddat$Estimated_fecundity) / sqrt(nrow(fecunddat))  / 1e6), 
  3)

```

***

## Exploratory Plots

### 1. egg volume ~ carapace width

```{r}
p1 <- ggplot(eggdat) +
  geom_point(aes(Carapace_width, new_vol, color = Unique_ID), alpha = 0.5, show.legend = FALSE) +
  facet_wrap(~Egg_stage)

p1
#plotly::ggplotly(p1)

```

### 2. Egg Volume ~ Egg Stage

```{r}
eggdat %>% 
  ggplot(aes(Egg_stage, new_vol)) + 
  geom_boxplot() +ylab("Volume mm^3")

```

***

# Estimated Fecundity Models {.tabset .tabset-fade .tabset-pills}

### Fecundity Range :

Estimated fecundity ranged from `r fecundity_min` - `r fecundity_max` million eggs.

With mean fecundity of `r fecundity_mean` +/- `r fecundity_se`  millions of eggs.

## Gaussian w/ identity link

```{r}
#Plot relationships with different link functions
#Gaussian
m1_g <- glm(Estimated_fecundity ~ Carapace_width, data = fecunddat)

fecunddat %>% 
  ggplot(aes(Carapace_width, Estimated_fecundity)) + 
  geom_point() + geom_smooth(method = "glm", 
                             formula = y~x,
                             method.args = list(family = gaussian(link = 'identity'))) +
  scale_y_continuous(labels = comma_format())

tidy(m1_g) %>% kable()

```

***

## Gaussian w/ log link

```{r}
#Log link
m1_log <- glm(Estimated_fecundity ~ Carapace_width, data = fecunddat, family = gaussian(link = "log"))

fecunddat %>% 
  ggplot(aes(Carapace_width, Estimated_fecundity)) + 
  geom_point(color = "gray20") + 
  geom_smooth(method = "glm", 
              formula = y~x,
              method.args = list(family = gaussian(link = 'log')),
              color = "black",
              linetype = 2) +
  scale_y_continuous(labels = comma_format()) +
  #theme_classic() +
  xlab("Carapace Width (mm)") +
  ylab("Estimated Fecundity (Number of Eggs)")

tidy(m1_log) %>% kable()

daic <- round(AIC(m1_g) - AIC(m1_log), 2)

```

Delta AIC for log-link - gaussian models = `r daic`

***

## Region intercepts (ANCOVA)

```{r fecundity ancova}

#Zack wants to see if regions matter, so lets add them in now
m1_ancova <- glm(Estimated_fecundity ~ Carapace_width + region, data = fecunddat, family = gaussian(link = "log"))
#summary(m1)

#Make a prediction dataset to plot them individually
xrange <- seq(min(fecunddat$Carapace_width), max(fecunddat$Carapace_width), by = .1)
# pred.df <- data.frame("Carapace_width" = rep(xrange, 4), 
#                       "region" = c(rep("FL",          length(xrange)),
#                                    rep("MS Bight",    length(xrange)),
#                                    rep("South Texas", length(xrange)),
#                                    rep("TX/LA",       length(xrange)))
# )
# 
# pred.df$fits <- predict(m1, newdata =  pred.df, type = "response")
# ms <- dplyr::filter(pred.df, region == "MS Bight")
# tx <- dplyr::filter(pred.df, region == "South Texas")
# la <- dplyr::filter(pred.df, region == "TX/LA")
# fl <- dplyr::filter(pred.df, region == "FL")

pred.df <- data.frame("Carapace_width" = rep(xrange, 4), 
                      "region" = c(rep("Florida",         length(xrange)),
                                   rep("MS Bight",        length(xrange)),
                                   rep("Texas",           length(xrange)),
                                   rep("Louisiana",       length(xrange)))
)

pred.df$fits <- predict(m1_ancova, newdata =  pred.df, type = "response")
ms <- dplyr::filter(pred.df, region == "MS Bight")
tx <- dplyr::filter(pred.df, region == "Texas")
la <- dplyr::filter(pred.df, region == "Louisiana")
fl <- dplyr::filter(pred.df, region == "Florida")

# #That was a yuge pain
# p1 <- ggplot() +
#   geom_point(data = fecunddat, aes(Carapace_width, Estimated_fecundity)) +
#   geom_line(data = fl, aes(Carapace_width, fits, color = "FL", group = 1)) +
#   geom_line(data = ms, aes(Carapace_width, fits, color = "Ms Bight", group = 1)) +
#   geom_line(data = tx, aes(Carapace_width, fits, color = "South Texas", group = 1)) +
#   geom_line(data = la, aes(Carapace_width, fits, color = "TX/LA", group = 1))  +
#   theme_classic() +
#   xlab("Carapace Width (mm)") +
#   ylab("Estimated Fecundity (Number of Eggs)")
# 
# p1

#That was a yuge pain
p1 <- ggplot() +
  geom_point(data = fecunddat, aes(Carapace_width, Estimated_fecundity)) +
  geom_smooth(data = fecunddat,
              aes(Carapace_width, Estimated_fecundity),
              method = "glm",
              formula = y~x,
              method.args = list(family = gaussian(link = 'log')),
              color = "black",
              linetype = 2) +
  scale_y_continuous(labels = comma_format()) +
  geom_line(data = fl, aes(Carapace_width, fits, color = "Florida", group = 1)) +
  geom_line(data = ms, aes(Carapace_width, fits, color = "MS Bight", group = 1)) +
  geom_line(data = tx, aes(Carapace_width, fits, color = "Texas", group = 1)) +
  geom_line(data = la, aes(Carapace_width, fits, color = "Louisiana", group = 1))  +
  theme_classic() + theme(legend.title = element_blank()) +
  xlab("Carapace Width (mm)") +
  ylab("Estimated Fecundity (Number of Eggs)")

p1

tidy(m1_ancova) %>% kable()

```

***

## Carapace width region interaction

If we include region in the model we get kind of a weird result, that being that carapace width is no longer significant:

```{r fecundity interaction}
m1_r <- glm(Estimated_fecundity ~ Carapace_width * region, 
          data = fecunddat,
          family = gaussian(link = "log"))

#and plot
fecunddat %>% 
  ggplot(aes(Carapace_width, Estimated_fecundity,)) + 
  geom_point(aes(color = region)) + 
  geom_smooth(method = "glm", 
              formula = y ~ x,
              method.args = list(family = gaussian(link = 'log')),
              color = "black",
              linetype = 2) +
  scale_y_continuous(labels = comma_format()) +
  facet_wrap(~region) +
  theme_classic() + theme(legend.title = element_blank()) +
  xlab("Carapace Width (mm)") +
  ylab("Estimated Fecundity (Number of Eggs)")

tidy(m1_r) %>% kable()

```

Which is confusing because its pretty clear on its own.

***

## Carapace Width & Season

Check Season as well just to be sure.

```{r}
m1_s <- glm(Estimated_fecundity ~ Carapace_width + Season, 
          data = fecunddat,
          family = gaussian(link = "log"))

#Prediction dataframe
pred.df <- data.frame("Carapace_width" = rep(xrange, 4), 
                      "Season" = c(rep("Summer", length(xrange)),
                                   rep("Fall",   length(xrange)))
)

pred.df$fits <- predict(m1_s, newdata =  pred.df, type = "response")
summer <- dplyr::filter(pred.df, Season == "Summer")
fall <-   dplyr::filter(pred.df, Season == "Fall")

#That was a yuge pain
p1 <- ggplot() +
  geom_point(data = fecunddat, aes(Carapace_width, Estimated_fecundity)) +
  geom_line(data = summer, aes(Carapace_width, fits, color = "Summer", group = 1)) +
  geom_line(data = fall,   aes(Carapace_width, fits, color = "Fall", group = 1)) + 
  scale_y_continuous(labels = comma_format()) +
  theme_classic() + theme(legend.title = element_blank()) +
  xlab("Carapace Width (mm)") +
  ylab("Estimated Fecundity (Number of Eggs)")

p1

tidy(m1_s) %>% kable()

```

## Model Comparison Table

```{r}
#Comparison Table
summ.table <- do.call(rbind, lapply(list(m1_g, m1_log, m1_ancova, m1_r), broom::glance))

table.cols <- c("df.residual", "deviance", "AIC")
reported.table <- summ.table[table.cols]
names(reported.table) <- c("Resid. Df", "Resid. Dev", "AIC")

reported.table[['dAIC']] <-  with(reported.table, AIC - min(AIC))
reported.table[['weight']] <- with(reported.table, exp(- 0.5 * dAIC) / sum(exp(- 0.5 * dAIC)))
reported.table$AIC <- NULL
reported.table$weight <- round(reported.table$weight, 2)
reported.table$dAIC <- round(reported.table$dAIC, 1)
row.names(reported.table) <- c("base : Gaussian", "base : Log-link", "region : no interaction", "region : interaction")

reported.table %>% 
  arrange(dAIC) %>% 
  kable()
```

# Egg size models {.tabset .tabset-fade .tabset-pills}

## Investigate Seasons  - individual eggs

(if we have fall egg sizes), we do **not** have enough (n = 5)

We also have repeated measures for measurements on each egg mass, so gonna have the eggmass/crabID as a random effect as they are likely correlated. Otherwise we could use the average egg measurements for each crab.

```{r}
p1 <- ggplot(eggdat) +
  geom_point(aes(Carapace_width, new_vol, color = Season), alpha = 0.5, show.legend = FALSE) +
  #facet_grid(Egg_stage ~ Season) +
  labs(
    x = "Carapace Width",
    y = "Egg Volume (mm^3)"
  )

p1
#plotly::ggplotly(p1)

```

***

## GLM - individual eggs

```{r}
m1 <- lm(new_vol ~ Carapace_width, data = eggdat)

ggplot(eggdat, aes(x = Carapace_width, y = new_vol)) +
  #facet_wrap(~Unique_ID, nrow = 5) +
  geom_point() +
  theme_classic() +
  geom_hline(yintercept = mean(eggdat$new_vol), linetype = 2, 
             aes(color = "Overall Mean")) + 
  geom_smooth(data = eggdat, 
              aes(x = Carapace_width, y = new_vol, group = 1), 
              method = "glm") +
  scale_color_manual(#name = "",
                     values = c("Overall Mean"     = "gray")) +
  labs(
    x = "Carapace Width (mm)",
    y = "Egg Volume",
    title = "Base Model"
    )

#Model Results
tidy(m1) %>% kable()

#qqplots
qqnorm(resid(m1))
qqline(resid(m1))
```

***

## GLM - eggmass means

```{r}
m1 <- lm(egg_volume ~ Carapace_width, data = fecunddat)

ggplot(fecunddat, aes(x = Carapace_width, y = egg_volume)) +
  #facet_wrap(~Unique_ID, nrow = 5) +
  geom_point() +
  theme_classic() +
  geom_hline(yintercept = mean(eggdat$new_vol), linetype = 2, 
             aes(color = "Overall Mean")) + 
  geom_smooth(data = eggdat, 
              aes(x = Carapace_width, y = new_vol, group = 1), 
              method = "glm") +
  scale_color_manual(#name = "",
                     values = c("Overall Mean"     = "gray")) +
  labs(
    x = "Carapace Width (mm)",
    y = "Egg Volume",
    title = "Base Model"
    )

#Model Results
tidy(m1) %>% kable()

```

***

## Mixed Model - individual eggs

```{r, warning = FALSE, message = FALSE}

#guide: http://bbolker.github.io/mixedmodels-misc/glmmFAQ.html#model-specification

# Carapace width plus unique id intercept
m2 <- lme4::lmer(new_vol ~ Carapace_width + (1 | Unique_ID), data = eggdat)

#summary(m2)

# # Confidence intervals
# confint(m2)

# # Random Effects
# lme4::ranef(m2)

#Variance Partitioning - From summary(m2)
re_var <- 2.904e-06
resid_var_total <- 2.904e-06 + 1.000e-06 #Total residual variance *AFTER* Fixed effects
re_varpart <- re_var / resid_var_total * 100

#Fixed Effects Plot
(fe_plot <- ggplot(eggdat, aes(x = Carapace_width, y = new_vol)) +
  #facet_wrap(~Unique_ID, nrow = 5) +
  geom_point() +
  theme_classic() +
  geom_hline(yintercept = mean(eggdat$new_vol), linetype = 2,
             aes(color = "Overall Mean")) + 
  geom_smooth(data = eggdat, 
              aes(x = Carapace_width, y = new_vol, group = 1), 
              method = "glm") +
  geom_point(data = cbind(eggdat, pred = predict(m2)), 
             aes(y = pred, color = "Individual Means"), 
             size = 2) +
  scale_color_manual(name = "",
                     values = c("Individual Means" = "red",
                                "Overall Mean"     = "gray")
                     ) +
  theme(legend.position = "right") +
  labs(
    x = "Carapace Width (mm)",
    y = "Egg Volume",
    title = "Overall Effect"
    )   
  )

#Model Results
tidy(m2) %>% kable() #not significant

# #Save this code for other MEM ****
# #Random Effects Plot - No line because single x value for each one
# (mm_plot <- ggplot(eggdat, aes(x = Carapace_width, y = new_vol)) +
#   facet_wrap(~Unique_ID, nrow = 5) +
#   geom_point() +
#   theme_classic() +
#   geom_line(data = cbind(eggdat, pred = predict(m2)), aes(y = pred)) +
#   theme(legend.position = "none") +
#   labs(
#     x = "Carapace Width (mm)",
#     y = "Egg Volume",
#     title = "Individual - Level Effect"
#     ) +
#     zero_margin
#   )

```

### Mixed Model Diagnostics

```{r}
# Diagnostics

#qqplots
qqnorm(resid(m2))
qqline(resid(m2))
```

```{r, fig.height=8}
# tt <- tidy(m2, effects="ran_vals")
# ggplot(tt,aes(level,estimate))+
#     geom_pointrange(aes(ymin=estimate-1.96*std.error,
#                         ymax=estimate+1.96*std.error))+
#     coord_flip()

# #Fitted vs. residuals + qq
# plot(m2)
# qqmath(m2)

# #group residuals
# plot(m2, Unique_ID ~ resid(.))

# #can also use augment from broom + ggplot
# aa <- broom.mixed::augment(m2)
# ggplot(aa, aes(Unique_ID, .resid)) +
#   geom_boxplot() + 
#   coord_flip()

# #Plot random effects to look for outliers:
# dotplot(ranef(m2, condVar=TRUE))

#or again with ggplot arranged in ascending order
tt <- tidy(m2, effects = "ran_vals") %>% 
  arrange(estimate) %>% 
  mutate(level = fct_inorder(level) #Order by ascending estimate
         # , 
         # significance = if_else(between(0, 
         #                               estimate - 1.96 * std.error, 
         #                               estimate + 1.96 * std.error),
         #                       "Not Significant",
         #                       "Significant")
         )

tt$significance <- NA

for (i in 1:nrow(tt)) {
  tt$significance[i] <- if_else(between(0, 
                                       tt$estimate[i] - 1.96 * tt$std.error[i],
                                       tt$estimate[i] + 1.96 * tt$std.error[i]),
                               "Not Significant",
                               "Significant")

}

ggplot(tt, aes(level, estimate))+
  geom_hline(yintercept = 0, linetype = 2, color = "navy", alpha = 0.5) +
  geom_pointrange(aes(ymin = estimate - 1.96 * std.error,
                      ymax = estimate + 1.96 * std.error,
                      alpha = significance)) + 
  scale_alpha_manual("", values = c("Significant" = 1,
                                    "Not Significant" = 0.3)) +

  coord_flip()

# #Use merTools to plot interval estimates
# library(merTools)
# merTools::plotREsim(REsim(m2))  # plot the interval estimates
```

### Mixed Model Inference

```{r}

####  Inference

#coefficient Plot
dwplot(m2) + geom_vline( xintercept = 0, lty = 2)

#P-values - 
library("lmerTest")
lmer2R <- lmer(new_vol ~ Carapace_width + (1 | Unique_ID), data = eggdat)
detach("package:lmerTest")

#Now get p=values with summary
#summary(lmer2R) #default uses the Satterthwaite approximation to the degrees of freedom. You can

#also use

#summary(lmer2R, ddf="Kenward-Roger") #for Kenward-roger which are more accurrate, take longer
#They are the same here

#Here are the type II (sequential) ANOVA results:
anova(lmer2R) #done withlmerTest gives you denominator DF
#anova(m2)     #regular lmer does not

#afex mixed can also be used to get the kenward-roger p-value
#afex::mixed(new_vol ~ Carapace_width + (1 | Unique_ID), data = eggdat)

#### Testing random effect
#If we want to test the random effect, we would in principle remove the random effect and test with anova, '

m1 <- glm(new_vol ~ Carapace_width, data = eggdat)
anova(m2, m1)

```

### Variance Explained By Random Effect

This is the variance explained after the fixed effects : `r round(re_varpart, 2)`

***

## Egg size ~ CW * region  - eggmass mean

Plot the overall relationship first

```{r}
# #Individual eggs
# eggdat %>% 
#   ggplot(aes(Carapace_width, new_vol)) +
#   geom_point(aes(color = region)) +
#   geom_smooth(method = "glm",
#               formula = y~x,
#               method.args = list(family = gaussian(link = 'identity'))) +
#   labs(
#     x = "Carapace Width",
#     y = "Egg Volume (mm^3)"
#   )

#mean eggmass values
fecunddat %>% 
  ggplot(aes(Carapace_width, egg_volume)) +
  geom_point(aes(color = region)) +
  geom_smooth(method = "glm",
              formula = y~x,
              method.args = list(family = gaussian(link = 'identity'))) +
  labs(
    x = "Carapace Width",
    y = "Egg Volume (mm^3)"
  )
```

Region seems to have a viable impact, but it may just look that way because of an interaction effect with egg size.

```{r}
# #individual eggs
# eggdat %>% #facet region and egg stage
#   ggplot(aes(Carapace_width, new_vol)) +
#   geom_point(aes(color = Egg_stage), show.legend = FALSE) +
#   facet_grid(region ~ Egg_stage) +
#   theme(axis.text.x = element_text(angle = 90)) +
#   labs(
#     x = "Carapace Width",
#     y = "Egg Volume (mm^3)"
#   )

fecunddat %>% #facet region and egg stage
  ggplot(aes(Carapace_width, egg_volume)) +
  geom_point(aes(color = Egg_stage), show.legend = FALSE) +
  facet_grid(region ~ Egg_stage) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(
    x = "Carapace Width",
    y = "Egg Volume (mm^3)"
  )

```

I don't think we will be able to tease this out with so many combinations with no data unless we re-bin into more general groups or drop regions like florida.

When you look at the interactions you get this mess:

```{r}
# m1 <- glm(new_vol ~ Carapace_width + Egg_stage * region, 
#           data = eggdat,
#           family = gaussian(link = "identity"))

m1 <- glm(egg_volume ~ Carapace_width + Egg_stage * region, 
          data = fecunddat,
          family = gaussian(link = "identity"))

tidy(m1) %>% kable()
```

***

## Egg Stage Bins

### Egg Volume ~ Egg Bins

```{r}

fecunddat %>% #facet region and egg stage
  ggplot(aes(Carapace_width, egg_volume)) +
  geom_point(aes(color = Egg_stage), show.legend = FALSE) +
  facet_grid(region ~ es_bins) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(
    x = "Carapace Width",
    y = "Egg Volume (mm^3)"
  )

fecunddat %>% 
  ggplot(aes(es_bins, egg_volume)) +
  geom_boxplot() +
  labs(
    x = "Egg Development Stage",
    y = "Egg Volume (mm^3)"
  )

anova1 <- lm(egg_volume ~ es_bins, data = fecunddat) 
summary(anova1)
# anova1 <- aov(egg_volume ~ es_bins, data = fecunddat) 
# tidy(anova1) %>% kable()

```

### Egg Volume ~ Region + Egg Bins

```{r}

fecunddat %>% 
  ggplot(aes(Carapace_width, egg_volume)) +
  geom_point(aes(color = es_bins)) +
  geom_smooth(method = "glm",
              formula = y~x,
              method.args = list(family = gaussian(link = 'identity'))) +
  facet_wrap(~region) +
  labs(
    x = "Carapace Width",
    y = "Egg Volume (mm^3)"
  ) + theme_minimal() + zero_margin

```

## Mean Egg Volume GLM - Final

 * Carapace Width     
 * Region (not enough sampling)      
 * Season

```{r}
ggplot(fecunddat, aes(Carapace_width, egg_volume)) +
  geom_point(aes(color = es_bins)) +
  geom_smooth(method = "glm") +
  facet_wrap(~es_bins)

m1 <- glm(egg_volume ~ Carapace_width * es_bins, data = fecunddat)
tidy(m1) %>% kable()

```

***

# Percent development Models

## Percent developing normally ~ molt stage

This one is nice and cut-and-dry, just an ANOVA to test the group differences.

```{r}
fecunddat %>% 
  ggplot(aes(molt_stage, Percent_fert)) +
  geom_boxplot(aes(color = molt_stage))

anova1 <- lm(Percent_fert ~ molt_stage, data = fecunddat)
summary(anova1)
#anova1 <- aov(Percent_fert ~ molt_stage, data = fecunddat)
#tidy(anova1) %>% kable()

```

Groups are unbalanced, and it throws a warning, but we're probably ok.

```{r}
fecunddat %>% count(molt_stage) %>% kable()
```

***

## Percent developing normally ~ Egg stage

```{r}
fecunddat %>% 
  ggplot(aes(fct_rev(es_bins), Percent_fert)) +
  #geom_jitter(aes(color = region), height = 0, width = 0.1) + 
  #geom_boxplot(aes(color = region)) + 
  geom_boxplot(aes(color = es_bins)) + 
  #coord_flip() +
  ylim(c(0,1)) +
  labs(
    x = "Egg Stage",
    y = "% Fertilized"
  )

anova1 <- lm(Percent_fert ~ es_bins, data = fecunddat)
summary(anova1)
#tidy(anova1) %>% kable()

```

***

## Percent developing normally ~ region

```{r}
fecunddat %>% 
  ggplot(aes(region, Percent_fert)) +
  #geom_jitter(aes(color = region), height = 0.1, width = 0.1)
  geom_boxplot(aes(color = region))

anova1 <- lm(Percent_fert ~ region, data = fecunddat)
summary(anova1)
#tidy(anova1) %>% kable()

```

***

## Percent Development - All

This is a two-way crossed ANOVA comparing percent fertilization between egg stage bins and region with interactions.

```{r}

fecunddat %>% 
  ggplot(aes(es_bins, Percent_fert)) +
  #geom_jitter(aes(color = region), height = 0.1, width = 0.1)
  geom_boxplot(aes(color = es_bins)) +
  facet_wrap(~region) +
  labs(caption = "Too much going on...")

fecunddat %>% 
  group_by(region, es_bins) %>% 
  summarise(n = n(),
            mean_fert = mean(Percent_fert))

# anova1 <- aov(Percent_fert ~ region * es_bins * molt_stage, data = fecunddat)
# tidy(anova1) %>% kable()

anova1 <- lm(Percent_fert ~ region * es_bins, data = fecunddat)
summary(anova1)
#tidy(anova1) %>% kable()

```

***