### 2020-01-05 2019_edits/R/offshore_analyses/crabs_per_hectare.Rmd

---
title: "Offshore Crabs/Hectare Modeling"
---
title: "Offshore Indices of Abundance for <br/> Blue Crabs"
author: "Adam A Kemberling"
date: "1/4/2020"
output: 
  html_document:
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

#packages
library(MASS)
library(here)
library(tidyverse)
library(knitr)
library(emmeans)
library(patchwork)
library(broom)
library(statmod)
library(sf)
conflicted::conflict_prefer("select", "dplyr")
conflicted::conflict_prefer("filter", "dplyr")

#Format ggplot for rest of document
theme_set(theme_bw() + theme(panel.grid.major = element_blank(), 
                             panel.grid.minor = element_blank(), 
                             panel.border = element_rect(fill = NA)))

`%not in%` <- purrr::negate(`%in%`)
```

```{r data}
seamap <- read_csv("~/Dropbox/SEAMAP_2019/data/offshore/seamap_cpue_2019.csv",
                   guess_max = 1e6, 
                   col_types = cols())

#Set up Season Column, using survey titles
seamap[str_detect(seamap$TITLE, "Fall"),"Season"] <- "Fall"
seamap[str_detect(seamap$TITLE, "Winter"),"Season"] <- "Winter"
seamap[str_detect(seamap$TITLE, "Summer"),"Season"] <- "Summer"
seamap[str_detect(seamap$TITLE, "Spring"),"Season"] <- "Spring"

# Make any setup changes
seamap <- seamap %>% 
  mutate(STAT_ZONE     = factor(STAT_ZONE),
         Crab_presence = ifelse(Sapidus_Catch > 0, 1, 0),
         Survey_Year   = as.numeric(Survey_Year),
         Season        = factor(Season, levels = c("Summer","Fall","Winter")),
         Survey_Month  = lubridate::month(Date),
         Survey_Month  = factor(Survey_Month),
         Survey_design = ifelse(Survey_Year < 2009, "Old Design", "Current Design"),
         Survey_design = ifelse(Survey_Year == 2008 & Season == "Fall", "Current Design", Survey_design),
         Survey_design = factor(Survey_design, levels = c("Old Design", "Current Design")),
         ####  More changes from DE
         Year_f             =  factor(Survey_Year),
         Month_f            =  factor(Survey_Month),
         Depth              =  (Start_Depth + End_Depth)/2,
         Temp_std           =  (Temp_B- mean(Temp_B))/ mean(Temp_B),
         Salinity_std       =  (Salinity_B- mean(Salinity_B))/ mean(Salinity_B),
         Depth_std          =  (Depth- mean(Depth))/ mean(Depth))

# Do all filtering
seamap <- seamap %>% 
  filter(Season            !=  "Winter",
         Survey_Year     %in%  c(2000:2018),
         Spd_kmh           !=  0,                    #these are infinite cpue
         Depth             <=  110)

# Catch only set
crabs <- seamap %>% filter(Sapidus_Catch > 0)

crabs %>% 
  mutate(STAT_ZONE = fct_rev(STAT_ZONE)) %>% 
  ggplot(aes(STAT_ZONE, Sapidus_Catch)) + 
  geom_boxplot(aes(group = STAT_ZONE)) + 
  labs(title    = "Blue Crab Catch by Stat Zone", 
       subtitle = "Stat zones ordered West to East - 2000 to 2018", 
       x        = "Statistical Zone",
       y        = "C. sapidus Single Station Catch") +
  facet_wrap(~Survey_design)

overall_mean <- mean(seamap$Sapidus_Catch, na.rm = T)
percent_zeros <- seamap %>% filter(Sapidus_Catch == 0) %>% count()
station_count <- seamap %>% count()
```