### 2020-02-10 2019_edits/R/offshore_analyses/final_analyses/Manuscript_Offshore_Modeling.Rmd
pulled manuscript edits out into their own markdown

---
title: "Offshore Modeling - Manuscript Model"
author: "Adam A. Kemberling"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

#packages
library(MASS)
library(here)
library(tidyverse)
library(knitr)
library(emmeans)
library(patchwork)
library(gridExtra)
library(broom)
library(statmod)
library(sf)
conflicted::conflict_prefer("select", "dplyr")
conflicted::conflict_prefer("filter", "dplyr")

#Format ggplot for rest of document
theme_set(theme_bw() + theme(panel.grid.major = element_blank(), 
                             panel.grid.minor = element_blank(), 
                             panel.border = element_rect(fill = NA),
                             axis.text = element_text(size = 12, color = "black"),
                             axis.title = element_text(size = 14, color = "black")))

`%not in%` <- purrr::negate(`%in%`)
```

# Manuscript Model

```{r load export}
seamap <- read_csv(str_c("~/Dropbox/SEAMAP_2019/data/offshore/", 
                        "seamap_manuscriptmods_full.csv"), 
                   col_types = cols()) %>% 
  mutate(year_f = factor(year_f),
         Season = factor(season, levels = c("Summer", "Fall")),
         region = factor(region, levels = c("Texas", "Louisiana", "MS Bight", "Florida"))
)

seamap_west <- read_csv(str_c("~/Dropbox/SEAMAP_2019/data/offshore/", 
                        "seamapwest_manuscriptmods_full.csv"), 
                        col_types = cols()) %>% 
  mutate(year_f = factor(year_f),
         Season = factor(season, levels = c("Summer", "Fall")),
         region = factor(region, levels = c("Texas", "Louisiana", "MS Bight", "Florida"))
)
```

## Model Dataset {.tabset}

### Model Data - West Only

```{r seamap west stats}
start_year <- min(as.numeric(as.character(seamap_west$survey_year)))
end_year <- max(as.numeric(as.character(seamap_west$survey_year)))
total_catch <- sum(seamap_west$sapidus_catch, na.rm = T)
overall_foa <- round(mean(seamap_west$crab_presence, na.rm = TRUE), 3) * 100
overall_mean <- round(mean(seamap_west$cpue_towspd, na.rm = T), 3)
percent_zeros <- seamap_west %>% filter(sapidus_catch == 0) %>% count()
station_count <- seamap_west %>% count()

```

 * Total Number of Stations = `r nrow(seamap_west)`

 * Starting in `r start_year` and continuing through `r end_year`

 * Total Sapidus Caught was `r total_catch`

 * Overall Frequency of Occurrence was `r overall_foa`

 * Overall Mean Catch Crabs/Hectare: `r overall_mean`

 * Stations with no catch: `r percent_zeros` / `r station_count` Stations **or** `r round((percent_zeros / station_count) * 100, 2)`%

### East + West

```{r seamap all stats}
start_year <- min(as.numeric(as.character(seamap$survey_year)))
end_year <- max(as.numeric(as.character(seamap$survey_year)))
total_catch <- sum(seamap$sapidus_catch, na.rm = T)
overall_foa <- round(mean(seamap$crab_presence, na.rm = TRUE), 3) * 100
overall_mean <- round(mean(seamap$cpue_towspd, na.rm = T), 3)
percent_zeros <- seamap %>% filter(sapidus_catch == 0) %>% count()
station_count <- seamap %>% count()

```

 * Total Number of Stations = `r nrow(seamap)`

 * Starting in `r start_year` and continuing through `r end_year`

 * Total Sapidus Caught was `r total_catch`

 * Overall Frequency of Occurrence was `r overall_foa`

 * Overall Mean Catch Crabs/Hectare: `r overall_mean`

 * Stations with no catch: `r percent_zeros` / `r station_count` Stations **or** `r round((percent_zeros / station_count) * 100, 2)`%

## Yearly Catch Plot {.tabset}

### Data Summary

*NOTE: Contains Data Across All Years and Regions (Including Florida)*

```{r}
year_sums <- seamap %>% 
  group_by(survey_year) %>% 
  summarise(Stations          = n(),
            `Missing Obs`     = sum(is.na(cpue_towspd)),
            `Mean Catch`      = mean(cpue_towspd, na.rm = T),
            `SD`              = sd(cpue_towspd, na.rm = T),
            `Var`             = var(cpue_towspd),
            `Overdispersed`   = ifelse(Var > `Mean Catch`, "yes", "no"),
            `Catch SE`        = sd(cpue_towspd, na.rm = T) / sqrt(Stations - `Missing Obs`),
            `% Occurrence`    = mean(crab_presence, na.rm = T),
            `% Occurrence SE` = sd(crab_presence, na.rm = T) / sqrt(Stations - `Missing Obs`)
            ) %>% 
  ungroup() %>% 
    mutate(Survey_Year        = factor(survey_year),
           catch_ymin         = `Mean Catch` - (2 * `Catch SE`),
           catch_ymax         = `Mean Catch` + (2 * `Catch SE`),
           catch_ymin         = ifelse(catch_ymin < 0, 0, catch_ymin),
           occurrence_ymin    = `% Occurrence` - (2 * `% Occurrence SE`),
           occurrence_ymax    = `% Occurrence` + (2 * `% Occurrence SE`),
           occurrence_ymin    = ifelse(occurrence_ymin < 0, 0, occurrence_ymin)
         )

year_sums %>% select(`Survey Year` = survey_year, Stations, `Missing Obs`, `% Occurrence`, `Mean Catch`, SD, `Overdispersed`) %>% knitr::kable()
```

### Annual Mean Catch

```{r}

year_plot <- year_sums %>% 
  mutate(survey_year = as.numeric(as.character(survey_year))) %>% 
  ggplot() +
    geom_col(aes(survey_year, `Mean Catch`)) +
    geom_errorbar(aes(x = survey_year, 
                      ymin = catch_ymin,
                      ymax = catch_ymax),
                  size = .25) +
  labs(x = NULL,
       y = "Mean Catch (C. sapidus / Hectare)")

year_plot
```

### Annual Frequency of Occurrence

```{r}
year_foa <- year_sums %>% 
  mutate(survey_year = as.numeric(as.character(survey_year))) %>% 
  ggplot() +
    geom_col(aes(survey_year, `% Occurrence`)) +
    # geom_errorbar(aes(x = survey_year, 
    #                   ymin = occurrence_ymin,
    #                   ymax = occurrence_ymax),
    #               size = .25) +
  scale_y_continuous(labels = scales::percent, limits = c(0,1)) + 
  labs(x = NULL, y = "Percent Occurrence")

year_foa
```

### Side x Side

```{r}

year_plot + year_foa
```

## Region - Season Plot {.tabset}

*NOTE: Contains Data Across All Years and Regions (Including Florida)*

*NOTE 2: Mean Catch (per hectare) and frequency of occurrence calculated for each year, in each region, for each season. The mean and standard error of those annual measurements are then taken*

### Factor Summary

```{r }
# rs_summary <- seamap %>% 
#   filter(is.na(region)      == FALSE) %>% 
#   group_by(region, season) %>% 
#   summarise(Stations          = n(),
#             `Missing Obs`     = sum(is.na(cpue_towspd)),
#             `Mean Catch`      = mean(cpue_towspd, na.rm = T),
#             `SD`              = sd(cpue_towspd, na.rm = T),
#             `Var`             = var(cpue_towspd),
#             `Overdispersed`   = ifelse(Var > `Mean Catch`, "yes", "no"),
#             `Catch SE`        = sd(cpue_towspd, na.rm = T) / sqrt(Stations - `Missing Obs`),
#             `% Occurrence`    = mean(crab_presence, na.rm = T),
#             `% Occurrence SE` = sd(crab_presence, na.rm = T) / sqrt(Stations - `Missing Obs`)) %>% 
#     ungroup() %>% 
#     mutate(Season             = factor(season, levels = c("Summer", "Fall")),
#            region             = factor(region, 
#                                        levels = c("Texas", "Louisiana", "MS Bight", "Florida")),
#            #region             = fct_rev(region),
#            catch_ymin         = `Mean Catch` - (2 * `Catch SE`),
#            catch_ymax         = `Mean Catch` + (2 * `Catch SE`),
#            catch_ymin         = ifelse(catch_ymin < 0, 0, catch_ymin),
#            occurrence_ymin    = `% Occurrence` - (2 * `% Occurrence SE`),
#            occurrence_ymax    = `% Occurrence` + (2 * `% Occurrence SE`),
#            occurrence_ymin    = ifelse(occurrence_ymin < 0, 0, occurrence_ymin)
#     )

# #Summary Table
# rs_summary %>%
#   select(Region = region, Season, Stations, `Missing Obs`, `% Occurrence`, `Mean Catch`, SD, Overdispersed) %>%
#   knitr::kable()

#Calaculate Annual average cpue and percent occurrence
rs_summary <- seamap %>% 
  group_by(survey_year, region, season) %>% 
  summarise(cpue = mean(cpue_towspd, na.rm = T),
            percent_occurrence = mean(crab_presence, na.rm = T)) %>% 
  group_by(region, season) %>% 
  summarise(nyears = n(),
            avg_annual_cpue = mean(cpue, na.rm = T),
            annual_cpue_se = sd(cpue, na.rm = T) / sqrt(nyears),
            avg_percent_occurrence = mean(percent_occurrence, na.rm = T),
            po_se = sd(percent_occurrence, na.rm = T) / sqrt(nyears)) %>% 
    ungroup() %>%
    mutate(Season             = factor(season, levels = c("Summer", "Fall")),
           region             = factor(region, levels = c("Texas", "Louisiana", 
                                                          "MS Bight", "Florida")),
           catch_ymin         = avg_annual_cpue - (2 * annual_cpue_se),
           catch_ymax         = avg_annual_cpue + (2 * annual_cpue_se),
           catch_ymin         = ifelse(catch_ymin < 0, 0, catch_ymin),
           occurrence_ymin    = avg_percent_occurrence - (2 * po_se),
           occurrence_ymax    = avg_percent_occurrence + (2 * po_se),
           occurrence_ymin    = ifelse(occurrence_ymin < 0, 0, occurrence_ymin),
           occurrence_ymax    = ifelse(occurrence_ymax > 1, 1, occurrence_ymax)
    )

```

### Mean Catch by Region and Season

```{r}

# Plot
rs_plot <- rs_summary %>% 
  ggplot(aes(region, avg_annual_cpue, fill = season)) +
    geom_col(position = "dodge") +
    geom_errorbar(aes(ymin = catch_ymin,
                      ymax = catch_ymax),
                  position = "dodge", 
                  size = .25) +
  scale_fill_manual(name = "", values = setNames(c("gray35", "gray50"), c("Summer", "Fall"))) +
  labs(x = NULL,
       y = "Mean Catch (C. sapidus / Hectare)") + 
  theme(
    legend.position = c(1, .95),
    legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(0, 6, 2, 2),
    legend.background = element_rect(fill = NA))

rs_plot
```

### Frequency of Occurrence

```{r}
rs_plot_2 <- rs_summary %>% 
  ggplot(aes(region, avg_percent_occurrence, fill = Season)) +
    geom_col(position = "dodge") +
    geom_errorbar(aes(ymin = occurrence_ymin,
                      ymax = occurrence_ymax),
                  position = "dodge",
                  size = .25) +
  scale_fill_manual(name = "", values = setNames(c("gray35", "gray50"), c("Summer", "Fall"))) +
  scale_y_continuous(labels = scales::percent, limits = c(0,1)) + 
  labs(x = NULL,
       y = "Percent Occurrence") + theme(
    legend.position = c(1, .95),
    legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(0, 6, 2, 2),
    legend.background = element_rect(fill = NA))

rs_plot_2
```

### Side x Side

```{r}
rs_plot + rs_plot_2

```

## Seasonal Catch Plot {.tabset}

*NOTE 2: Mean Catch (per hectare) and frequency of occurrence calculated for each year, for each season. The mean and standard error of those annual measurements are then taken*

### Season Factor Summary

```{r}
# season_summary <- seamap %>% 
#   group_by(survey_year, season) %>% 
#   summarise(Stations          = n(),
#             `Missing Obs`     = sum(is.na(cpue_towspd)),
#             `Mean Catch`      = mean(cpue_towspd, na.rm = T),
#             `SD`              = sd(cpue_towspd, na.rm = T),
#             `Var`             = var(cpue_towspd),
#             `Overdispersed`   = ifelse(Var > `Mean Catch`, "yes", "no"),
#             `Catch SE`        = sd(cpue_towspd, na.rm = T) / sqrt(Stations - `Missing Obs`),
#             `% Occurrence`    = mean(crab_presence, na.rm = T),
#             `% Occurrence SE` = sd(crab_presence, na.rm = T) / sqrt(Stations - `Missing Obs`)) %>% 
#     ungroup() %>% 
#     mutate(survey_year        = factor(survey_year),
#            # StatZone         = factor(StatZone),
#            # StatZone         = fct_rev(StatZone),
#            catch_ymin         = `Mean Catch` - (2 * `Catch SE`),
#            catch_ymax         = `Mean Catch` + (2 * `Catch SE`),
#            catch_ymin         = ifelse(catch_ymin < 0, 0, catch_ymin),
#            occurrence_ymin    = `% Occurrence` - (2 * `% Occurrence SE`),
#            occurrence_ymax    = `% Occurrence` + (2 * `% Occurrence SE`),
#            occurrence_ymin    = ifelse(occurrence_ymin < 0, 0, occurrence_ymin)
#          )
# 
# season_summary %>% select(`Survey Year` = survey_year, Season = season, Stations, `Missing Obs`, `% Occurrence`, `Mean Catch`, SD, Overdispersed) %>% knitr::kable()

#Calaculate Annual average cpue and percent occurrence
season_summary <- seamap %>% 
  group_by(survey_year, season) %>% 
  summarise(cpue = mean(cpue_towspd, na.rm = T),
            percent_occurrence = mean(crab_presence, na.rm = T)) %>% 
  group_by(season) %>% 
  summarise(nyears = n(),
            avg_annual_cpue = mean(cpue, na.rm = T),
            annual_cpue_se = sd(cpue, na.rm = T) / sqrt(nyears),
            avg_percent_occurrence = mean(percent_occurrence, na.rm = T),
            po_se = sd(percent_occurrence, na.rm = T) / sqrt(nyears)) %>% 
    ungroup() %>%
    mutate(Season             = factor(season, levels = c("Summer", "Fall")),
           catch_ymin         = avg_annual_cpue - (2 * annual_cpue_se),
           catch_ymax         = avg_annual_cpue + (2 * annual_cpue_se),
           catch_ymin         = ifelse(catch_ymin < 0, 0, catch_ymin),
           occurrence_ymin    = avg_percent_occurrence - (2 * po_se),
           occurrence_ymax    = avg_percent_occurrence + (2 * po_se),
           occurrence_ymin    = ifelse(occurrence_ymin < 0, 0, occurrence_ymin),
           occurrence_ymax    = ifelse(occurrence_ymax > 1, 1, occurrence_ymax)
    )

```

### Seasonal Catch Plot

```{r}

season_plot <- season_summary %>% 
  ggplot(aes(season, avg_annual_cpue, fill = season)) +
    geom_col() +
    geom_errorbar(aes(ymin = catch_ymin,
                      ymax = catch_ymax),
                  size = .25) +
  scale_fill_manual(name = "", values = setNames(c("gray35", "gray50"), c("Summer", "Fall"))) +
  labs(x = NULL,
       y = "Mean Catch (C. sapidus / Hectare)") + theme(
    legend.position = c(1, .95),
    legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(0, 6, 2, 2),
    legend.background = element_rect(fill = NA))

season_plot

```

### Seasonal Frequency of Occurrence

```{r}

season_plot_2 <- season_summary %>% 
  ggplot(aes(season, avg_percent_occurrence, fill = season)) +
    geom_col(position = "dodge") +
    geom_errorbar(aes(ymin = occurrence_ymin,
                      ymax = occurrence_ymax),
                  size = .25) +
  scale_fill_manual(name = "", values = setNames(c("gray35", "gray50"), c("Summer", "Fall"))) +
  scale_y_continuous(labels = scales::percent, limits = c(0,1)) + 
  labs(x = NULL,
       y = "Percent Occurrence") + theme(
    legend.position = c(1, .95),
    legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(0, 6, 2, 2),
    legend.background = element_rect(fill = NA))

season_plot_2
```

### Side x Side

```{r}
season_plot + season_plot_2
```

## Candidate Models {.tabset}

### Model Selection using AIC

```{r}

#Candidate 1 - no interactions
candidate_1 <- MASS:::glm.nb(sapidus_catch ~ year_f + region + season +   #Additive covariates
            #year_f * region + 
            offset(log(area_hectares)),
            data = seamap_west)

#Candidate 2 - year * region
candidate_2 <- MASS:::glm.nb(sapidus_catch ~ year_f + region + season +   #Additive covariates
            year_f * region + 
            offset(log(area_hectares)),
            data = seamap_west)

#Candidate 3 - year * season
candidate_3 <- MASS:::glm.nb(sapidus_catch ~ year_f + region + season +   #Additive covariates
            year_f * season + 
            offset(log(area_hectares)),
            data = seamap_west)

#Candidate 4 - region * season
candidate_4 <- MASS:::glm.nb(sapidus_catch ~ year_f + region + season +   #Additive covariates
            region * season + 
            offset(log(area_hectares)),
            data = seamap_west)

#Candidate 5 - year * region + region * season
candidate_5 <- MASS:::glm.nb(sapidus_catch ~ year_f + region + season +   #Additive covariates
            year_f * region + region * season +
            offset(log(area_hectares)),
            data = seamap_west)

model_list <- list(
  "no interactions" = candidate_1,
  "year * region" = candidate_2,
  "year * season" = candidate_3,
  "region * season" = candidate_4,
  "all interactions" = candidate_5
)

model_aic <- map(model_list,AIC)
model_aic <- bind_rows(
  imap(model_aic, function(x,y) {
    aic_df <- data.frame("Formulation" = y, AIC = x) }
    )
  )
model_aic %>% mutate(`Delta AIC` = AIC - min(AIC)) %>% knitr::kable()

best_mod <- candidate_2
rm(candidate_1, candidate_2, candidate_3, candidate_4, candidate_5)
```

### Predictor Significance

```{r}

#Drop1 method
#drop1(mod1a, test = "F")     # F-test
#drop1(mod1a, test = "Chisq") # Chisq LRT - no scope

#Name what you want to drop
lrt_y  <- drop1(best_mod, scope = ~ year_f, test = "Chisq")
lrt_r  <- drop1(best_mod, scope = ~ region, test = "Chisq")
lrt_s  <- drop1(best_mod, scope = ~ season, test = "Chisq")
lrt_yr <- drop1(best_mod, scope = ~ year_f*region, test = "Chisq")

bind_rows(
  list(
    lrt_y  %>% as.data.frame() %>% 
      rownames_to_column(var = "Predictor"),
    lrt_r  %>% as.data.frame() %>% 
      rownames_to_column(var = "Predictor") %>% filter(Predictor == "region"),
    lrt_s  %>% as.data.frame() %>% 
      rownames_to_column(var = "Predictor") %>% filter(Predictor == "season"),
    lrt_yr  %>% as.data.frame() %>% 
      rownames_to_column(var = "Predictor") %>% filter(Predictor == "year_f:region")
)) %>% kable()

#broom::glance(best_mod) %>% kable()
#Percent deviance explained
perc_dev <- (best_mod$null.deviance - best_mod$deviance) / best_mod$null.deviance
perc_dev <- round(perc_dev * 100,2)
```

*Percent Deviance Explained by Best Candidate Model:* `r perc_dev`

### Model Summary

Statistic reported below is the Wald Chi-Square z-score.

```{r}
#summary(best_mod) #to see what test statistic is used
broom::tidy(best_mod) %>% kable()
```

### Interaction Plot

```{r}
emmip(best_mod, region ~ year_f, type = "response", offset = 0) + labs(x = NULL, caption = "Interaction Plot for year * region factors")
```

## Catch per Hectare Plot

```{r}

#Estimated Marginal Means with an offset
#parameter in the call of offset = 0:
best_mod_dat <- as.data.frame(emmeans(best_mod, 
                                  specs = "year_f", 
                                  by    = c("region", "season"),
                                  type  = "response", 
                                  offset = 0))
best_mod_dat$CV <- best_mod_dat$SE/best_mod_dat$response

#Drop Infinite credibility intervals
best_mod_dat <- best_mod_dat %>% 
  mutate(asymp.UCL = ifelse(asymp.UCL == Inf, NA, asymp.UCL))

# Catch Plot
p <- best_mod_dat %>% 
  mutate(year_n = as.numeric(as.character(year_f))) %>% 
  ggplot(aes(year_n, response)) +
      geom_line(aes(year_n, response, color = season)) +
      geom_point(aes(year_n, response, color = season, shape = season)) +
      scale_color_grey(start = 0.2, end = 0.6) + 
      labs(
        x = NULL, 
        y = "Crabs per Hectare",
        title = "Model Estimates and Yearly observed means") +
      facet_grid(region ~ .) +
  theme(
    legend.position = c(1, .98),
    legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(0, 6, 2, 2),
    legend.background = element_rect(fill = NA),
    legend.title = element_blank())
p

```

## Raster Plots  {.tabset}

```{r spatial data setup}

library(stars)
library(sf)
library(rnaturalearth)

#State and Country Bounds
usmap <- ne_states(country = "united states of america") %>% 
  st_as_sf() %>% 
  filter(region == "South")

mex <- ne_states(country = "mexico") %>% 
  st_as_sf()

## Statistical Zones

#Load the simplified nmfs grid
szones <- read_sf("~/Dropbox/SEAMAP_2019/data/gis_files/nmfs_grids_clipped_rough.shp", 
                  crs = 4326) #%>% st_simplify()

#Custom grid for rasterize
test_bbox <- st_bbox(c(xmin = -97.5, xmax = -81,
                       ymin = 24.3, ymax = 30.5),
                     crs = st_crs(4326))

num_x <- seq(from = -97.4, to = -81.3, by = 0.2)
num_y <- seq(from = 24.4, to = 30.4, by = 0.2)

####  Aggregation  ####
#Aggregating the seamap model into bins
seamap_sf_agg <- seamap %>% 
  mutate(lon_bin = round(start_long, 1),
         lat_bin = round(start_lat, 1)) %>% 
  group_by(lon_bin, lat_bin) %>% 
  summarise(n_stations = n(),
            total_catch = sum(sapidus_catch, na.rm = T),
            avg_cpue = mean(cpue_towspd)) %>% 
  st_as_sf(coords = c("lon_bin", "lat_bin"), crs = 4326, remove = FALSE)

####  Total Catch  ####

#Convert to Raster and plot 
catch_ras <- st_rasterize(
  seamap_sf_agg["total_catch"],
  template = st_as_stars(
      test_bbox,
      #st_bbox(seamap_sf), 
      values = NA_real_,
      nx = length(num_x),
      ny = length(num_y))
  )

#plot(catch_ras, axes = T, main = "Total Sapidus Catch - All Years 2000+")

####  Mean CPUE  ####

cpue_ras <- st_rasterize(
  seamap_sf_agg["avg_cpue"],
  template = st_as_stars(
      test_bbox,
      #st_bbox(seamap_sf), 
      values = NA_real_,
      nx = length(num_x),
      ny = length(num_y))
  )

#plot(cpue_ras, axes = T, main = " Mean Catch/Hectare - All Years 2000+")

####  Sampling Effort  ####
effort_ras <- st_rasterize(
  seamap_sf_agg["n_stations"],
  template = st_as_stars(
      test_bbox,
      #st_bbox(seamap_sf), 
      values = NA_real_,
      nx = length(num_x),
      ny = length(num_y))
  )

```

### Statistical Zones
```{r}

#Numbering for statzones
sz_df <- tribble(

~"SZ", ~"long",  ~"lat",
21,    -96.5,     26.5,
20,    -96.7,  27.5,
19,    -95.75,  28.25,
18,    -94.5,   28.8,
17,    -93.5,   28.9,
16,    -92.5,   28.85,
15,    -91.5,   28.8,
14,    -90.5,   28.75,
13,    -89.75,  28.9,
12,    -89.25,  29.65,
11,    -88.5,   29.75,
10,    -87.5,   29.80,
 9,    -86.5,   29.80,
 8,    -85.5,   29.35,
 7,    -84.20,   29.45,
 6,    -83.8,   28.50,
 5,    -83.5,   27.55,
 4,    -83.35,   26.75,
 3,    -83.35,   25.75,
 2,    -84.20,   24.55,
 2,    -81.55,   24.28,
 1,    -82.5,   24.28

) %>% mutate(SZ = as.character(SZ))

ggplot() +
  geom_sf(data = szones, fill = NA) +
  geom_sf(data = usmap) +
  geom_sf(data = mex) +
  geom_text(data = sz_df, aes(long, lat, label = SZ), size = 3.5) +
  coord_sf(xlim = c(-98, -81), ylim = c(24, 31)) +
  labs(x = NULL, y = NULL)  +
  scale_y_continuous(breaks = c(25,30))
```

### Sampling Effort

```{r}
####  Sampling Effort  ####
ggplot() +
  geom_stars(data = effort_ras) +
  geom_sf(data = usmap) +
  geom_sf(data = mex) +
  scale_fill_distiller(palette = "Spectral", na.value = "transparent") +
  coord_sf(xlim = c(-98, -81), ylim = c(24, 31)) +
  scale_y_continuous(breaks = c(25,30)) +
  labs(x = NULL, y = NULL) +
  guides(fill = guide_colorbar(title = "Total Sampling Effort (Stations)",
                               title.hjust = 0.5, 
                               title.position = "top",
                               barwidth = 10)) +
  theme(legend.position = "bottom")
```

### Catch per Hectare

```{r}

####  CPUE  ####
ggplot() +
  geom_stars(data = cpue_ras) +
  geom_sf(data = usmap) +
  geom_sf(data = mex) +
  scale_fill_distiller(palette = "Spectral", na.value = "transparent") +
  coord_sf(xlim = c(-98, -81), ylim = c(24, 31)) +
  scale_y_continuous(breaks = c(25,30)) +
  labs(x = NULL, y = NULL) +
  guides(fill = guide_colorbar(title = "Average Catch/Hectare",
                               title.hjust = 0.5, 
                               title.position = "top",
                               barwidth = 10)) +
  theme(legend.position = "bottom")

```

### Total Catch

```{r}
####  Total Catch  ####
ggplot() +
  geom_stars(data = catch_ras) +
  geom_sf(data = usmap) +
  geom_sf(data = mex) +
  scale_fill_distiller(palette = "Spectral", na.value = "transparent") +
  coord_sf(xlim = c(-98, -81), ylim = c(24, 31)) +
  scale_y_continuous(breaks = c(25,30)) +
  labs(x = NULL, y = NULL) +
  guides(fill = guide_colorbar(title = "Total C. sapidus Caught",
                               title.hjust = 0.5, 
                               title.position = "top",
                               barwidth = 10)) +
  theme(legend.position = "bottom")

```