### 2020-02-15 2019_edits/R/offshore_analyses/final_analyses/Manuscript_Offshore_Modeling.Rmd
figure changes

---
title: "Offshore Modeling - Manuscript Model"
author: "Adam A. Kemberling"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
  code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

#packages
library(MASS)
library(here)
library(tidyverse)
library(knitr)
library(emmeans)
library(patchwork)
library(gridExtra)
library(broom)
library(statmod)
library(sf)
conflicted::conflict_prefer("select", "dplyr")
conflicted::conflict_prefer("filter", "dplyr")

#Format ggplot for rest of document
theme_set(theme_bw() + 
          theme(panel.grid.major = element_blank(), 
            panel.grid.minor = element_blank(), 
            panel.border = element_rect(fill = NA),
            axis.text = element_text(size = 12, color = "black"),
            axis.title = element_text(size = 12, color = "black"),
            legend.text = element_text(size = 12, color = "black"),
            legend.background = element_rect(fill = NA))
          )

`%not in%` <- purrr::negate(`%in%`)
```

# Manuscript Model

```{r load data}
seamap <- read_csv(str_c("~/Dropbox/SEAMAP_2019/data/offshore/", 
                        "seamap_manuscriptmods_full.csv"), 
                   col_types = cols()) %>% 
  mutate(year_f = factor(year_f),
         Season = factor(season, levels = c("Summer", "Fall")),
         region = ifelse(region == "Louisiana", "LA West", region),
         region = factor(region, levels = c("Texas", "LA West", "MS Bight", "Florida"))
)

seamap_west <- read_csv(str_c("~/Dropbox/SEAMAP_2019/data/offshore/", 
                        "seamapwest_manuscriptmods_full.csv"), 
                        col_types = cols()) %>% 
  mutate(year_f = factor(year_f),
         Season = factor(season, levels = c("Summer", "Fall")),
         region = ifelse(region == "Louisiana", "LA West", region),
         region = factor(region, levels = c("Texas", "LA West", "MS Bight", "Florida"))
)
```

## Model Dataset {.tabset}

### Model Data - West Only

```{r seamap west stats}
start_year <- min(as.numeric(as.character(seamap_west$survey_year)))
end_year <- max(as.numeric(as.character(seamap_west$survey_year)))
total_catch <- sum(seamap_west$sapidus_catch, na.rm = T)
overall_foa <- round(mean(seamap_west$crab_presence, na.rm = TRUE), 3) * 100
overall_mean <- round(mean(seamap_west$cpue_towspd, na.rm = T), 3)
percent_zeros <- seamap_west %>% filter(sapidus_catch == 0) %>% count()
station_count <- seamap_west %>% count()

```

 * Total Number of Stations = `r nrow(seamap_west)`

 * Starting in `r start_year` and continuing through `r end_year`

 * Total Sapidus Caught was `r total_catch`

 * Overall Frequency of Occurrence was `r overall_foa`

 * Overall Mean Catch Crabs/Hectare: `r overall_mean`

 * Stations with no catch: `r percent_zeros` / `r station_count` Stations **or** `r round((percent_zeros / station_count) * 100, 2)`%

### East + West

```{r seamap all stats}
start_year <- min(as.numeric(as.character(seamap$survey_year)))
end_year <- max(as.numeric(as.character(seamap$survey_year)))
total_catch <- sum(seamap$sapidus_catch, na.rm = T)
overall_foa <- round(mean(seamap$crab_presence, na.rm = TRUE), 3) * 100
overall_mean <- round(mean(seamap$cpue_towspd, na.rm = T), 3)
percent_zeros <- seamap %>% filter(sapidus_catch == 0) %>% count()
station_count <- seamap %>% count()

```

 * Total Number of Stations = `r nrow(seamap)`

 * Starting in `r start_year` and continuing through `r end_year`

 * Total Sapidus Caught was `r total_catch`

 * Overall Frequency of Occurrence was `r overall_foa`

 * Overall Mean Catch Crabs/Hectare: `r overall_mean`

 * Stations with no catch: `r percent_zeros` / `r station_count` Stations **or** `r round((percent_zeros / station_count) * 100, 2)`%

## Annual Catch and Occurrence {.tabset}

### Data Summary

*NOTE: Contains Data Only from Western-Gulf Subset*

```{r}
year_sums <- seamap_west %>% 
  group_by(survey_year) %>% 
  summarise(Stations          = n(),
            `Missing Obs`     = sum(is.na(cpue_towspd)),
            `Mean Catch`      = mean(cpue_towspd, na.rm = T),
            `SD`              = sd(cpue_towspd, na.rm = T),
            `Var`             = var(cpue_towspd),
            `Overdispersed`   = ifelse(Var > `Mean Catch`, "yes", "no"),
            `Catch SE`        = sd(cpue_towspd, na.rm = T) / sqrt(Stations - `Missing Obs`),
            `% Occurrence`    = mean(crab_presence, na.rm = T),
            `% Occurrence SE` = sd(crab_presence, na.rm = T) / sqrt(Stations - `Missing Obs`)
            ) %>% 
  ungroup() %>% 
    mutate(Survey_Year        = factor(survey_year),
           catch_ymin         = `Mean Catch` - (2 * `Catch SE`),
           catch_ymax         = `Mean Catch` + (2 * `Catch SE`),
           catch_ymin         = ifelse(catch_ymin < 0, 0, catch_ymin),
           occurrence_ymin    = `% Occurrence` - (2 * `% Occurrence SE`),
           occurrence_ymax    = `% Occurrence` + (2 * `% Occurrence SE`),
           occurrence_ymin    = ifelse(occurrence_ymin < 0, 0, occurrence_ymin)
         )

year_sums %>% 
  select(`Survey Year` = survey_year, Stations, `Missing Obs`, 
         `% Occurrence`, `Mean Catch`, SD, `Overdispersed`) %>% knitr::kable()
```

### Annual Mean Catch

```{r, fig.height=4}

west_mean <- round(mean(seamap_west$cpue_towspd, na.rm = T), 3)

year_plot <- year_sums %>% 
  mutate(survey_year = as.numeric(as.character(survey_year))) %>% 
  ggplot() +
    geom_col(aes(survey_year, `Mean Catch`), fill = "gray35") +
    geom_errorbar(aes(x = survey_year, 
                      ymin = catch_ymin,
                      ymax = catch_ymax),
                  size = .25) +
  geom_hline(yintercept = west_mean, 
             linetype = "dashed") +
  labs(x = NULL,
       y = expression(paste("Mean CPUE (crabs ", ha^{-1},")"))
  )

year_plot
```

### Annual Frequency of Occurrence

```{r, fig.height=4}
west_foa <- round(mean(seamap_west$crab_presence, na.rm = TRUE), 3)

year_foa <- year_sums %>% 
  mutate(survey_year = as.numeric(as.character(survey_year))) %>% 
  ggplot() +
    geom_col(aes(survey_year, `% Occurrence`)) +
    geom_hline(yintercept = west_foa, 
               linetype = "dashed") +
    scale_y_continuous(labels = scales::percent, limits = c(0,1)) + 
    labs(x = NULL, y = "Percent occurrence")

year_foa
```

### Side x Side

```{r, fig.height=4}

year_plot + year_foa

```

## Region - Season Interactions {.tabset}

*NOTE: Contains Data Only from Western-Gulf Subset*

*NOTE 2: Mean Catch (per hectare) calculated as the average crabs/hectare from each tow within a region-season grouping. Percent occurrence is calculated separately as the average annual percent occurrence for those same groups.*

### Factor Summary

```{r }

#Mean CPUE for the regions within each season
rs_summary <- seamap_west %>%
  filter(is.na(region)      == FALSE) %>%
  group_by(region, season) %>%
  summarise(Stations          = n(),
            `Missing Obs`     = sum(is.na(cpue_towspd)),
            `Mean Catch`      = mean(cpue_towspd, na.rm = T),
            `SD`              = sd(cpue_towspd, na.rm = T),
            `Var`             = var(cpue_towspd),
            `Overdispersed`   = ifelse(Var > `Mean Catch`, "yes", "no"),
            `Catch SE`        = sd(cpue_towspd, na.rm = T) / sqrt(Stations - `Missing Obs`)) %>%
    ungroup() %>%
    mutate(Season             = factor(season, levels = c("Summer", "Fall")),
           region             = factor(region,
                                       levels = c("Texas", "LA West", "MS Bight", "Florida")),
           #region             = fct_rev(region),
           catch_ymin         = `Mean Catch` - (2 * `Catch SE`),
           catch_ymax         = `Mean Catch` + (2 * `Catch SE`),
           catch_ymin         = ifelse(catch_ymin < 0, 0, catch_ymin)
    )

#Summary Table
rs_summary %>%
  select(Region = region, Season, Stations, `Missing Obs`, `Mean Catch`, SD, Overdispersed) %>%
  knitr::kable()
```

Average frequency of occurrence for each region and season is calculated in two steps. In the first step the frequency of occurrence for each region and for both survey seasons is calculated for every year. In the second step those annual means are averaged to get an overall measure for each region for each season, and a measure of error around them is calculated to represent the inter-annual variation in the frequency of occurrence.

```{r}

#Calaculate Annual average percent occurrence
rs_foa <- seamap_west %>% 
  group_by(survey_year, region, season) %>% 
  summarise(percent_occurrence = mean(crab_presence, na.rm = T)) %>% 
  group_by(region, season) %>% 
  summarise(nyears = n(),
            avg_percent_occurrence = mean(percent_occurrence, na.rm = T),
            po_se = sd(percent_occurrence, na.rm = T) / sqrt(nyears)) %>% 
    ungroup() %>%
    mutate(Season             = factor(season, levels = c("Summer", "Fall")),
           region             = factor(region, levels = c("Texas", "LA West", 
                                                          "MS Bight", "Florida")),
           occurrence_ymin    = avg_percent_occurrence - (2 * po_se),
           occurrence_ymax    = avg_percent_occurrence + (2 * po_se),
           occurrence_ymin    = ifelse(occurrence_ymin < 0, 0, occurrence_ymin),
           occurrence_ymax    = ifelse(occurrence_ymax > 1, 1, occurrence_ymax)
    )

rs_foa %>% 
  select(Region = region, Season, `N Years` = nyears, 
         `% Occurrence` = avg_percent_occurrence, `% Occurrence SE` = po_se) %>%
  knitr::kable()

```

### Mean Catch by Region and Season

```{r, fig.height=4}

# Plot
rs_plot <- rs_summary %>% 
  ggplot(aes(region, `Mean Catch`, fill = Season)) +
    geom_col(position = "dodge") +
    geom_errorbar(aes(ymin = catch_ymin,
                      ymax = catch_ymax),
                  position = "dodge", 
                  size = .25) +
  scale_fill_manual(name = "", values = setNames(c("gray35", "gray70"), c("Summer", "Fall"))) +
  labs(x = NULL,
       y = expression(paste("Mean CPUE (crabs ", ha^{-1},")"))) + 
  theme(
    legend.position = c(1, .95),
    legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(0, 6, 2, 2))

rs_plot
```

### Frequency of Occurrence

```{r, fig.height=4}
rs_plot_2 <- rs_foa %>% 
  ggplot(aes(region, avg_percent_occurrence, fill = Season)) +
    geom_col(position = "dodge") +
    geom_errorbar(aes(ymin = occurrence_ymin,
                      ymax = occurrence_ymax),
                  position = "dodge",
                  size = .25) +
  scale_fill_manual(name = "", values = setNames(c("gray35", "gray70"), c("Summer", "Fall"))) +
  scale_y_continuous(labels = scales::percent, limits = c(0,1)) + 
  labs(x = NULL,
       y = "Percent occurrence") + 
  theme(
    legend.position = c(1, .95),
    legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(0, 6, 2, 2))

rs_plot_2
```

### Side x Side

```{r fig.height=4}
rs_plot <- rs_plot + theme(legend.position = "none")
rs_plot + rs_plot_2

```

## Candidate Models {.tabset}

### Model Selection using AIC

```{r}

#Candidate 1 - no interactions
candidate_1 <- MASS:::glm.nb(sapidus_catch ~ year_f + region + season +   #Additive covariates
            #year_f * region + 
            offset(log(area_hectares)),
            data = seamap_west)

#Candidate 2 - year * region
candidate_2 <- MASS:::glm.nb(sapidus_catch ~ year_f + region + season +   #Additive covariates
            year_f * region + 
            offset(log(area_hectares)),
            data = seamap_west)

#Candidate 3 - year * season
candidate_3 <- MASS:::glm.nb(sapidus_catch ~ year_f + region + season +   #Additive covariates
            year_f * season + 
            offset(log(area_hectares)),
            data = seamap_west)

#Candidate 4 - region * season
candidate_4 <- MASS:::glm.nb(sapidus_catch ~ year_f + region + season +   #Additive covariates
            region * season + 
            offset(log(area_hectares)),
            data = seamap_west)

#Candidate 5 - year * region + region * season
candidate_5 <- MASS:::glm.nb(sapidus_catch ~ year_f + region + season +   #Additive covariates
            year_f * region + region * season +
            offset(log(area_hectares)),
            data = seamap_west)

model_list <- list(
  "no interactions" = candidate_1,
  "year * region" = candidate_2,
  "year * season" = candidate_3,
  "region * season" = candidate_4,
  "all interactions" = candidate_5
)

model_aic <- map(model_list,AIC)
model_aic <- bind_rows(
  imap(model_aic, function(x,y) {
    aic_df <- data.frame("Formulation" = y, AIC = x) }
    )
  )
model_aic %>% mutate(`Delta AIC` = AIC - min(AIC)) %>% knitr::kable()

best_mod <- candidate_2
rm(candidate_1, candidate_2, candidate_3, candidate_4, candidate_5)
```

### Predictor Significance

```{r}

#Drop1 method
#drop1(mod1a, test = "F")     # F-test
#drop1(mod1a, test = "Chisq") # Chisq LRT - no scope

#Name what you want to drop
lrt_y  <- drop1(best_mod, scope = ~ year_f, test = "Chisq")
lrt_r  <- drop1(best_mod, scope = ~ region, test = "Chisq")
lrt_s  <- drop1(best_mod, scope = ~ season, test = "Chisq")
lrt_yr <- drop1(best_mod, scope = ~ year_f*region, test = "Chisq")

bind_rows(
  list(
    lrt_y  %>% as.data.frame() %>% 
      rownames_to_column(var = "Predictor"),
    lrt_r  %>% as.data.frame() %>% 
      rownames_to_column(var = "Predictor") %>% filter(Predictor == "region"),
    lrt_s  %>% as.data.frame() %>% 
      rownames_to_column(var = "Predictor") %>% filter(Predictor == "season"),
    lrt_yr  %>% as.data.frame() %>% 
      rownames_to_column(var = "Predictor") %>% filter(Predictor == "year_f:region")
)) %>% kable()

#broom::glance(best_mod) %>% kable()
#Percent deviance explained
perc_dev <- (best_mod$null.deviance - best_mod$deviance) / best_mod$null.deviance
perc_dev <- round(perc_dev * 100,2)
```

*Percent Deviance Explained by Best Candidate Model:* `r perc_dev`

### Model Summary

Statistic reported below is the Wald Chi-Square z-score.

```{r}
#summary(best_mod) #to see what test statistic is used
broom::tidy(best_mod) %>% kable()
```

### Interaction Plot

```{r}
emmip(best_mod, region ~ year_f, type = "response", offset = 0) + labs(x = NULL, caption = "Interaction Plot for year * region factors")
```

## Catch per Hectare Plot {.tabset .tabset-pills}

### Seasonal Prediction Differences

```{r}

#Estimated Marginal Means with an offset
#parameter in the call of offset = 0:
best_mod_dat <- as.data.frame(emmeans(best_mod, 
                                  specs = "year_f", 
                                  by    = c("region", "season"),
                                  type  = "response", 
                                  offset = 0))
best_mod_dat$CV <- best_mod_dat$SE/best_mod_dat$response

#Drop Infinite credibility intervals
best_mod_dat <- best_mod_dat %>% 
  mutate(asymp.UCL = ifelse(asymp.UCL == Inf, NA, asymp.UCL),
         season = factor(season, levels = c("Summer", "Fall")),
         region = factor(region, levels = c("Texas", "LA West", "MS Bight")))

# Catch Plot
p <- best_mod_dat %>% 
  mutate(year_n = as.numeric(as.character(year_f))) %>% 
  ggplot(aes(year_n, response)) +
      geom_line(aes(year_n, response, color = season)) +
      geom_point(aes(year_n, response, color = season, shape = season)) +
      scale_color_grey(start = 0.35, end = 0.7) + 
      labs(
        x = NULL, 
        y = expression(paste("Mean CPUE (crabs ", ha^{-1},")"))) +
      facet_grid(region ~ .) +
  theme(
    legend.position = c(1, .98),
    legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(0, 6, 2, 2),
    legend.background = element_rect(fill = NA),
    legend.title = element_blank())
p

```

### Observed vs. Predicted Data

```{r}
#Observed data
mean_obs <- mean(seamap_west$cpue_towspd)
obs_timeline <- seamap_west %>% 
  group_by(survey_year, region, season) %>% 
  summarise(Nobs = n(),
            PercentPos = mean(crab_presence),
            Mean_CPUE = mean(cpue_towspd)) %>% 
  ungroup() %>% 
  mutate(season = factor(season, levels = c("Summer", "Fall")),
         region = factor(region, levels = c("Texas", "LA West", "MS Bight")),
         Index_Obs = Mean_CPUE/mean_obs)

#Overall group means
panel_means <- seamap_west %>% 
  group_by(region, season) %>% 
  summarise(Mean_CPUE = mean(cpue_towspd))

#ggplot of standardized index of abundance
p <- best_mod_dat %>% 
  mutate(year_n = as.numeric(as.character(year_f))) %>% 
  ggplot(aes(year_n, response)) +
    geom_hline(data = panel_means, 
               aes(yintercept = Mean_CPUE), alpha = 0.5) +
    # geom_ribbon(aes(year_n, 
    #                 ymin = asymp.LCL, 
    #                 ymax = asymp.UCL), 
    #             fill = "gray80") +
    geom_line(aes(x = year_n, 
                  y = response, 
                  group = 1, 
                  col = "Predicted")) +
    geom_point(aes(x = year_n, 
                   y = response, 
                   col = "Predicted")) +
    # Observed Data
    geom_line(data = obs_timeline, 
            aes(x = survey_year, 
                y = Mean_CPUE, 
                group = region, 
                col = "Observed"), linetype = 2) +
    geom_point(data = obs_timeline, 
               aes(x = survey_year, 
                   y = Mean_CPUE, 
                   group = region, 
                   col = "Observed")) +
    #Theme changes
    #scale_color_grey(start = 0.35, end = 0.7) + 
    scale_color_manual(values = c("gray35", "royalblue")) + 
    labs(
      x = NULL, 
      y = expression(paste("Mean CPUE (crabs ", ha^{-1},")"))) +
  theme(legend.position = c(1, .98),
    legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(0, 6, 2, 2),
    legend.background = element_rect(fill = NA),
    legend.title = element_blank()) +
  facet_grid(season~region)
p

```

## Raster Plots  {.tabset}

```{r spatial data setup}

library(stars)
library(sf)
library(rnaturalearth)

#State and Country Bounds
usmap <- ne_states(country = "united states of america") %>% 
  st_as_sf() %>% 
  filter(region == "South")

mex <- ne_states(country = "mexico") %>% 
  st_as_sf()

## Statistical Zones

#Load the simplified nmfs grid
szones <- read_sf("~/Dropbox/SEAMAP_2019/data/gis_files/nmfs_grids_clipped_rough.shp", 
                  crs = 4326) #%>% st_simplify()

#Custom grid for rasterize
test_bbox <- st_bbox(c(xmin = -97.5, xmax = -81,
                       ymin = 24.3, ymax = 30.5),
                     crs = st_crs(4326))

num_x <- seq(from = -97.4, to = -81.3, by = 0.2)
num_y <- seq(from = 24.4, to = 30.4, by = 0.2)

####  Aggregation  ####
#Aggregating the seamap model into bins
seamap_sf_agg <- seamap %>% 
  mutate(lon_bin = round(start_long, 1),
         lat_bin = round(start_lat, 1)) %>% 
  group_by(lon_bin, lat_bin) %>% 
  summarise(n_stations = n(),
            total_catch = sum(sapidus_catch, na.rm = T),
            avg_cpue = mean(cpue_towspd)) %>% 
  st_as_sf(coords = c("lon_bin", "lat_bin"), crs = 4326, remove = FALSE)

####  Total Catch  ####

#Convert to Raster and plot 
catch_ras <- st_rasterize(
  seamap_sf_agg["total_catch"],
  template = st_as_stars(
      test_bbox,
      #st_bbox(seamap_sf), 
      values = NA_real_,
      nx = length(num_x),
      ny = length(num_y))
  )

#plot(catch_ras, axes = T, main = "Total Sapidus Catch - All Years 2000+")

####  Mean CPUE  ####

cpue_ras <- st_rasterize(
  seamap_sf_agg["avg_cpue"],
  template = st_as_stars(
      test_bbox,
      #st_bbox(seamap_sf), 
      values = NA_real_,
      nx = length(num_x),
      ny = length(num_y))
  )

#plot(cpue_ras, axes = T, main = " Mean Catch/Hectare - All Years 2000+")

####  Sampling Effort  ####
effort_ras <- st_rasterize(
  seamap_sf_agg["n_stations"],
  template = st_as_stars(
      test_bbox,
      #st_bbox(seamap_sf), 
      values = NA_real_,
      nx = length(num_x),
      ny = length(num_y))
  )

```

### Statistical Zones
```{r}

#Numbering for statzones
sz_df <- tribble(

~"SZ", ~"long",  ~"lat",
21,    -96.5,     26.5,
20,    -96.7,  27.5,
19,    -95.75,  28.25,
18,    -94.5,   28.8,
17,    -93.5,   28.9,
16,    -92.5,   28.85,
15,    -91.5,   28.8,
14,    -90.5,   28.75,
13,    -89.75,  28.9,
12,    -89.25,  29.65,
11,    -88.5,   29.75,
10,    -87.5,   29.80,
 9,    -86.5,   29.80,
 8,    -85.5,   29.35,
 7,    -84.20,   29.45,
 6,    -83.8,   28.50,
 5,    -83.5,   27.55,
 4,    -83.35,   26.75,
 3,    -83.35,   25.75,
 2,    -84.20,   24.55,
 2,    -81.55,   24.28,
 1,    -82.5,   24.28

) %>% mutate(SZ = as.character(SZ))

ggplot() +
  geom_sf(data = szones, fill = NA) +
  geom_sf(data = usmap) +
  geom_sf(data = mex) +
  geom_text(data = sz_df, aes(long, lat, label = SZ), size = 3.5) +
  coord_sf(xlim = c(-98, -81), ylim = c(24, 31)) +
  labs(x = NULL, y = NULL)  +
  scale_y_continuous(breaks = c(25,30))
```

### Sampling Effort

```{r}
####  Sampling Effort  ####
ggplot() +
  geom_stars(data = effort_ras) +
  geom_sf(data = usmap) +
  geom_sf(data = mex) +
  scale_fill_distiller(palette = "Spectral", na.value = "transparent") +
  coord_sf(xlim = c(-98, -81), ylim = c(24, 31)) +
  scale_y_continuous(breaks = c(25,30)) +
  labs(x = NULL, y = NULL) +
  guides(fill = guide_colorbar(title = "Total Sampling Effort (Stations)",
                               title.hjust = 0.5, 
                               title.position = "top",
                               barwidth = 10)) +
  theme(legend.position = "bottom", 
        legend.text = element_text(size  = 12), 
        legend.title = element_text(size = 12))
```

### Catch per Hectare

```{r}

####  CPUE  ####
ggplot() +
  geom_stars(data = cpue_ras) +
  geom_sf(data = usmap) +
  geom_sf(data = mex) +
  scale_fill_distiller(palette = "Spectral", na.value = "transparent") +
  coord_sf(xlim = c(-98, -81), ylim = c(24, 31)) +
  scale_y_continuous(breaks = c(25,30)) +
  labs(x = NULL, y = NULL) +
  guides(fill = guide_colorbar(title = "Average Catch/Hectare",
                               title.hjust = 0.5, 
                               title.position = "top",
                               barwidth = 10)) +
  theme(legend.position = "bottom", 
        legend.text = element_text(size  = 12), 
        legend.title = element_text(size = 12))

```

### Total Catch

```{r}
####  Total Catch  ####
ggplot() +
  geom_stars(data = catch_ras) +
  geom_sf(data = usmap) +
  geom_sf(data = mex) +
  scale_fill_distiller(palette = "Spectral", na.value = "transparent") +
  coord_sf(xlim = c(-98, -81), ylim = c(24, 31)) +
  scale_y_continuous(breaks = c(25,30)) +
  labs(x = NULL, y = NULL) +
  guides(fill = guide_colorbar(title = "Total C. sapidus Caught",
                               title.hjust = 0.5, 
                               title.position = "top",
                               barwidth = 10)) +
  theme(legend.position = "bottom", 
        legend.text = element_text(size  = 12), 
        legend.title = element_text(size = 12))

```